High-Fidelity Counterfactual Explanations through Conformal Prediction

### 2023-08-09 notebooks/mnist.qmd
revision of rebuttals

```{julia}
function _plot_eccco_mnist(
    x::Union{AbstractArray, Int}=x_factual, target::Int=target;
    λ=[0.1,0.25,0.25],
    temp=0.1,
    plt_order = ["MLP", "MLP Ensemble", "JEM", "JEM Ensemble"],
    opt = nothing,
    rng::Union{Int,AbstractRNG}=1234,
    T::Int = 100,
    use_class_loss::Bool = true,
    model_dict=model_dict,
    wide::Bool = false,
    img_height::Int = img_height,
    plot_factual::Bool = false,
    kwrgs...,
)

@@

    # Generate counterfactuals using ECCCo generator:
    eccco_generator = ECCCoGenerator(
        λ=λ, 
        temp=temp, 
        opt=opt,
        use_class_loss=use_class_loss,
        nsamples=10,
        nmin=10,
    )

    ces = Dict()
    for (mod_name, mod) in model_dict
        ce = generate_counterfactual(
            x_fact, target, counterfactual_data, mod, eccco_generator; 
            decision_threshold=γ, max_iter=T,
            initialization=:identity,
            converge_when=:generator_conditions,
        )
        ces[mod_name] = ce
    end
    _plt_order = map(x -> findall(collect(keys(model_dict)) .== x)[1], plt_order)

@@

    return plt, eccco_generator, ces
end
```

@@

```{julia}
_plt_order = [
    "MLP", 
    "MLP Ensemble", 
    "Large Ensemble (n=50)", 
    "LeNet-5",
    # "LeNet-5 Ensemble",
    # "RobustNet",
    "JEM", 
    "JEM Ensemble",
]
plt_additional_models, _, _ces_ = _plot_eccco_mnist(
    plt_order = _plt_order,
    model_dict=large_model_dict,
    wide = true,
    plot_factual = true,
    img_height = 150,
)
display(plt_additional_models)
savefig(plt_additional_models, joinpath(output_images_path, "mnist_eccco_additional.png"))
```

@@

>>>>

```{julia}
function _plot_eccco_mnist(
    x::Union{AbstractArray, Int}=x_factual, target::Int=target;
    λ=[0.1,0.25,0.25],
    temp=0.1,
    plt_order = ["MLP", "MLP Ensemble", "JEM", "JEM Ensemble"],
    opt = nothing,
    rng::Union{Int,AbstractRNG}=1234,
    T::Int = 100,
    use_class_loss::Bool = true,
    model_dict=model_dict,
    wide::Bool = false,
    img_height::Int = img_height,
    plot_factual::Bool = false,
    generator::Union{Nothing,CounterfactualExplanations.AbstractGenerator}=nothing,
    test_data::Bool = false,
    kwrgs...,
)

@@

    if isnothing(generator)
        # Generate counterfactuals using ECCCo generator:
        generator = ECCCoGenerator(
            λ=λ, 
            temp=temp, 
            opt=opt,
            use_class_loss=use_class_loss,
            nsamples=10,
            nmin=10,
        )
    end

    if test_data
        data = load_mnist_test()
    else
        data = counterfactual_data
    end

    ces = Dict()
    for (mod_name, mod) in model_dict
        ce = generate_counterfactual(
            x_fact, target, data, mod, generator; 
            decision_threshold=γ, max_iter=T,
            initialization=:identity,
            converge_when=:generator_conditions,
        )
        ces[mod_name] = ce
    end
    _plt_order = map(x -> findall(collect(keys(model_dict)) .== x)[1], plt_order)

@@

    return plt, generator, ces
end
```

@@

```{julia}
_plt_order = [
    "MLP", 
    "MLP Ensemble", 
    "Large Ensemble (n=50)", 
    "LeNet-5",
    # "LeNet-5 Ensemble",
    # "RobustNet",
    "JEM", 
    "JEM Ensemble",
]
plt_additional_models, _, _ces_ = _plot_eccco_mnist(
    λ = [0.125,0.25,0.25],
    plt_order = _plt_order,
    model_dict=large_model_dict,
    wide = true,
    plot_factual = true,
    img_height = 150,
)
display(plt_additional_models)
# savefig(plt_additional_models, joinpath(output_images_path, "mnist_eccco_additional.png"))
```

@@

```{julia}
λ = [0.1,0.25,0.25]

wachter = WachterGenerator(
    λ=λ[1], 
    opt=eccco_generator.opt
)

combos = [
    (9,7),
    # (9,7),
    # (9,7),
    (6,0),
    # (6,0),
    # (6,0),
    (4,9),
    # (4,9),
    # (4,9),
    (8,3),
    # (8,3),
    # (8,3),
    (5,6),
    # (5,6),
    # (5,6),
]

plts_eccco = []
plts_wachter = []
ces_eccco = []
ces_wachter = []
rngs = []

for (factual, target) in combos

    rng = rand(1:10000)

    # ECCCo:
    plt, _, ces = _plot_eccco_mnist(
        factual, target; 
        λ = λ,
        plt_order = _plt_order,
        model_dict = large_model_dict,
        wide = true,
        plot_factual = true,
        rng = rng,
        img_height = 150
    )
    display(plt)
    push!(plts_eccco, plt)
    push!(ces_eccco, reduce(hcat, target_probs.(values(ces))))

    # Wachter:
    plt, _, ces = _plot_eccco_mnist(
        factual, target; 
        plt_order = _plt_order,
        model_dict = large_model_dict,
        wide = true,
        plot_factual = true,
        rng = rng,
        img_height = 150,
        generator = wachter,
    )
    display(plt)
    push!(plts_wachter, plt)
    push!(ces_wachter, reduce(hcat, target_probs.(values(ces))))

    push!(rngs, rng)

end
```
